---
title: "Lab 2 Practice"
author: "Donghyung Lee"
date: '2019-09-05'
output: 
  workflowr::wflow_html:
    toc: false
    code_folding: show
editor_options:
  chunk_output_type: console
---

<br />
**An Intro to R data programming** <br />

**Base R tools:** <br />
* classes   <br />
* numeric summaries <br />
* basic plots <br />
<br />
**New R tools:** <br />
* tidyverse (is a collection of R packages) <br />
* ggplot2 package: advanced graphics <br />
* dplyr package: data manipulation, working with data frames <br />
<br />

### Part 1: Object Type
```{r read_data}
icecream <- read.table("data/icecream.txt")
dim(icecream)
copier <- read.table("data/CH01PR20.txt")
dim(copier)
```

**Understand Dataframe**
```{r dataframe}
class(icecream)
class(copier)
head(icecream)
head(copier)
colnames(copier)=c("minutes","number")
head(copier)

#alternative way
copier <- setNames(copier,c("minutes","number"))
head(copier)

dim(icecream)
names(icecream)
class(icecream$video)
head(iris)
class(iris$Species)
class(iris$Sepal.Length)
```

**Class type can be changed**
```{r classtype}
class(icecream$video)
x=as.numeric(icecream$video)
class(x)
```

**Check the difference of the following to different types**
```{r typediff}
x
y=as.factor(x)
class(y)
y
z=as.character(x)
class(z)
z
```

**Change the variable ice_cream to factor**
```{r factor}
icecream$ice_cream
class(icecream$ice_cream)
icecream$ice_cream=as.factor(icecream$ice_cream)
icecream$ice_cream
class(icecream$ice_cream)
```

### Part 2: Exploratory Data Analysis

**(1) Numerical summaries:** <br />
**-mean, median, five number summary, standard deviation, IQR, correlation, etc.** <br />

**Traditional R**
```{r tradR}
mean(icecream$video)
median(icecream$video)
sd(icecream$video)
var(icecream$video)
summary(icecream$video)
IQR(icecream$video)
cor(copier$minutes,copier$number)

summary(icecream)
summary(copier)
```

**Advanced summary statistics (tidyverse)** <br />
“tibbles” instead of R’s traditional data.frame. 
Tibbles are data frames, but they tweak some older behaviours to make life a little easier. 

**Install and load R package "dplyr"**
```{r dplyr}
#install.packages("dplyr")
library(dplyr)
```

**pipe** <br />
Use %>%(pipe). You can read it as a series of imperative statements: group, then summarise, then filter. A good way to pronounce %>% when reading code is “then”.Behind the scenes, x %>% f(y) turns into f(x, y), and x %>% f(y) %>% g(z) turns into g(f(x, y), z) and so on. You can use the pipe to rewrite multiple operations in 
a way that you can read left-to-right, top-to-bottom.

**dplyr verbs** <br />
'filter','arrange','mutate','summarise','group_by'

**filter: select cases based on their values**
```{r filter}
head(icecream)
icecream <- as_tibble(icecream)
icecream

icecream %>% filter(female==0)
icecream %>% filter(female==1, video<50)

iris %>% filter(Species=="setosa", Sepal.Width>4)
iris %>% filter(Species=="versicolor", Petal.Length<4)

# Question: filter iris dataset for Species equal to "setosa" or "virginica"
```

**arrange: reorder cases**
```{r arrange}
icecream %>% arrange(video) # order 'video' column in ascending order
icecream %>% arrange(desc(puzzle)) # order 'puzzle' column in descending order

as_tibble(iris) %>% arrange(Petal.Length)
as_tibble(iris) %>% arrange(desc(Sepal.Length))

# Question: 1) filter iris dataset for Species equal to "setosa" and 2) sort  in descending order of Sepal.Width  
```

**mutate: add new variables that are functions of existing variables** 
```{r mutate}
icecream_new <- icecream %>% mutate(puzzle100 = puzzle*100)
icecream_new

# Question: 1) filter icecream dataset for ice_cream equal to 1, 2) create video1000 (video*1000) column, 3) sort in descending order of video1000, 4) assign the dataset to icecream_new2
```

**summarise: condense multiple values to a single value** 
```{r summarise}
icecream %>% summarise(Mean_video=mean(video), SD_video=sd(video), SD_median=median(video))
```

**group_by: break down a dataset into specified groups of rows**
```{r group_by}
puzzle.summary <- icecream %>% group_by(ice_cream) %>% summarise(Mean=mean(puzzle),
            Variance=var(puzzle))%>%as.data.frame()

iris %>% group_by(Species) %>% summarise(Mean=mean(Sepal.Length), Median=median(Sepal.Length), Variance=var(Sepal.Length))

# Question: 1) group by Species 2) calculate mean, median, var, min, max of Sepal.Length for each group 3) sort data in descending order of mean 3) convert to a data frame  4) assign the output to "iris_new"
```

**Average and standard deviation by icecream flavor** <br />
```{r pipe}
icecream %>% group_by(ice_cream) %>% 
  summarise(Mean=mean(puzzle),Variance=var(puzzle))
puzzle.summary <-icecream %>% group_by(ice_cream) %>% 
  summarise(Mean=mean(puzzle), Variance=var(puzzle) )
puzzle.summary
class(puzzle.summary)

puzzle.summary <- icecream %>% group_by(ice_cream) %>% 
  summarise(Mean=mean(puzzle),
            Variance=var(puzzle) )%>%as.data.frame()
puzzle.summary
class(puzzle.summary)
```
