---
title: "Lab Assignment 2 (Due Oct 8 Tuesday)"
author: "Donghyung Lee"
date: "`r Sys.Date()`"
output: 
  workflowr::wflow_html:
    code_folding: show
editor_options:
  chunk_output_type: console
---

Upload a pdf file or word file in Canvas generated using R markdown. You should clearly label the question number, include the r code, the output and any necessary explanation in your file. The plots should be made using ggplot2 package.

#### Load CalCOFI data using the following R codes:
```{r prep}
cofi <- read.table("https://raw.githubusercontent.com/dleelab/STA463_563_Fall2019/master/data/calcofi_500.csv", header=TRUE, sep = ",")
head(cofi)
```

#### Question 1 (1pt) Descriptive statistics for all variables in the data.
```{r Q1}
summary(cofi)
```

#### Question 2 (1pt) A density plot for one of the variables.
```{r Q2}
library(ggplot2)
ggplot(cofi, aes(x=sal)) + geom_density()

library(ggplot2)
ggplot(cofi, aes(x=temp)) + geom_density()

library(ggplot2)
ggplot(cofi, aes(x=depth)) + geom_density()
```

#### Question 3 (1pt) A scatter plot of salinity and log10(depth+1), use "salinity" as X-axis.
```{r Q3}
ggplot(cofi, aes(x=sal, y=log10(depth+1))) + geom_point() + theme_classic()
```

#### Question 4 (1pt) A correlation matrix of salinity and log10(depth+1).
```{r Q4}
cofi$log10depth <- log10(cofi$depth+1)
cor(cofi[,c(1,4)])
```

#### Question 5 (1pt) Use the lm() function in R, fit a simple linear regression using "salinity" as response and log10(depth+1) as predictor. Report the value of b0 and b1 in fitted simple linear regression equation.
```{r Q5}
cofi.fit <- lm(sal~log10depth, data=cofi)
summary(cofi.fit)
```

#### Question 6 (1pt) Verify the sum of the residuals equals to 0.
```{r Q6}
sum(cofi.fit$residuals)
```

#### Question 7 (1pt) Make a density plot for the residuals.
```{r Q7}
ggplot(cofi.fit, aes(x=cofi.fit$residuals)) + geom_density() 
```

#### Question 8 (3pt) Instead of using lm(), use R code to calculate b0 and b1 based on the formula we discussed in class. Verify the values match the results using the lm() function.
```{r Q8}
Sxx <- sum((cofi$log10depth-mean(cofi$log10depth))^2)
Sxx
Sxy <- sum((cofi$log10depth-mean(cofi$log10depth))*(cofi$sal-mean(cofi$sal)))
Sxy
b1 <- Sxy/Sxx
b1
b0 <- mean(cofi$sal)-b1*mean(cofi$log10depth)
b0
```
