---
title: "Simultaneous CI/PIs and RTTO Models"
author: "Donghyung Lee"
date: "`r Sys.Date()`"
output: 
  workflowr::wflow_html:
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Simultaneous CI/PIs

#### **Fit a simple linear regression model, estimate the expected number of minutes spent when there are 3,5 and 7 copiers to be serviced, respectively. Use a 90\% family confidence coefficient, calculate the simultaneous confidence interval, based on Working-Hotelling procedure and Bonferroni respectively.**

#### (a) W-H Method.
```{r}
data0120=read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01PR20.txt")
colnames(data0120)=c("minutes","copiers_number")
fit=lm(minutes~copiers_number,data=data0120)

alphaF=1-0.9
n=nrow(data0120)
W=sqrt(2*qf((1-alphaF),2,(n-2)))#Or W=sqrt(2*qf((1-alphaF),length(fit$coefficients),fit$df.residual))
Yh_hat=predict(fit,newdata=data.frame(copiers_number=c(3,5,7)),data=data0120)
MSE=sum((fit$residuals)^2)/(n-2)
xbar=mean(data0120$copiers_number)
Sxx=sum((data0120$copiers_number-xbar)^2)
s_Yh_hat=sqrt(MSE*(1/n+(c(3,5,7)-xbar)^2/Sxx))
simulCI=cbind(Yh_hat,Yh_hat-W*s_Yh_hat,Yh_hat+W*s_Yh_hat)
colnames(simulCI)=c("fitted","lower","upper")
rownames(simulCI)=c("3","5","7")
simulCI
```
The 90% simultaneous CI using W-H method is as above.

#### (b) Bonferroni method.
```{r}
g=3
alpha=alphaF/g
predict(fit,new=data.frame(copiers_number=c(3,5,7)),interval = "conf",level=1-alpha,data=data0120)
```
The 90% simultaneous CI using Bonferroni method is as above.


#### Discussion.
**From result above, which simultaneous confidence interval calculation method do you prefer? Why? (Therefore, in practice, before compute the simultaneous CI or PI, you would compare the multiplier first, and use whichever method that yields a smaller multiplier)**

Compare the results above, I would prefer the Bonferroni method. Because it provides a shorter confidence interval.
Note:
Therefore, before we compute the simultaneous CIs (PIs), we will compare the multipliers of different methods, then do the calculation. Here, in this example, multiplier for Bonferroni method is slightly smaller than the W-H method.
```{r}
t=qt((1-alphaF/g/2),(n-2))
W
t
```

### Regression through the origin.
**View the simple linear regression fit, and recall the discussion we had in class, adjust your model and fit a regression through the origin model (RTTO) instead. Obtain the estimated regression function.**
```{r}
summary(fit)
fit2=lm(minutes~copiers_number-1,data=data0120)
summary(fit2)
```
The estimated regression function is $\hat{Y_i}=14.9472X_i$, (or estimated minutes=14.9472* copiers_number)

#### Check the residuals.
**Check $\sum_{i=1}^{n}e_i\neq0$ and  $\sum_{i=1}^{n}e_iX_i=0$.**
```{r}
res=fit2$residuals
sum(res)
sum(data0120$copiers_number*res)
```
Based on the calculation above, $\sum_{i=1}^{n}e_i=-5.86$ is not equal to 0 and $\sum_{i=1}^{n}e_iX_i=-1.46*10^{-13}$, almost equal to 0.


#### Confidence interval.
**Estimate $\beta_1$ with a 95\% confidence interval.**
```{r}
confint(fit2,,level=0.95)
```
The 95% confidence interval for $\beta_1$ is (14.49,15.40).

#### Prediction interval.
**Predict the service time on a call in which six copiers are serviced. Use a 98\% confidence level.**
```{r}
predict(fit2,new=data.frame(copiers_number=6),interval="prediction",level=0.98,data=data0120)
```
The 98% prediction interval for the service time on a call in which 6 copiers are serviced is (68.15,111.22)

#### Simultaneous prediction interval.
**Estimate the number of minutes spent when there are 2,3,4,5 copiers to be serviced. Use a 95\% family confidence coefficient and use Bonferroni method. **
```{r}
alphaF=1-0.95
g=4
alpha=alphaF/g
predict(fit2,new=data.frame(copiers_number=c(2,3,4,5)),interval = "pred",level=1-alpha,data=data0120)
```

The 95% simultaneous PI for the number of minutes spent when there're 2,3,4,5 copiers to be serviced is as above.
