---
title: "Simple Linear Regression Model"
author: "Donghyung Lee"
date: "`r Sys.Date()`"
output: 
  workflowr::wflow_html:
    code_folding: show
editor_options:
  chunk_output_type: console
---

### Load data
```{r prep}
library(ggplot2)
copier <- read.table("data/CH01PR20.txt")
dim(copier)
colnames(copier)=c("minutes","number")
```

### Simple Linear Regression
```{r slr, eval=TRUE, echo=TRUE}
copier.fit <- lm(minutes~number,data=copier)
copier.fit
class(copier.fit)
str(copier.fit)
summary(copier.fit)
```
### Fitted regression line (traditional R functions)
```{r plot}
plot(minutes~number, data=copier)
abline(-0.5802,15.0352)#by definition of the line abline(intercept, slope)

#The following are alternative ways to draw the fitted regression line.
lines(copier$number,-0.5802+15.0352*copier$number)#other way

abline(copier.fit)#simple way
```

### Fitted regression line (ggplot2)
```{r ggplot}
fitted(copier.fit)
copier$fitted=fitted(copier.fit)
p <- ggplot(copier,aes(x=number, y=minutes)) + 
  geom_point()
p

p <- p + geom_line(aes(x=number,y=fitted))
p
```

### Results of lm()
Verify the lm() gives us the correct calculation result <br />

1. Residual standard error=sqrt(MSE)=sqrt(SSE/n-2), n-2 is the degree of freedom. We can verify this later. <br />
2. It gives a five number summary of the residuals, mean of residuals is? It's 0. We proved this during class. <br />
3. Since residuals are estimates of errors, under normal error SLR, 
we expect the residuals are also normally distributed. Are the residuals
normally distribued? <br />
<br />
Question 1. How to get the residuals? <br />
Question 2. How to get the estimated values? Are there other measures already calculated?


### Attributes of fitted model
```{r attributes}
attributes(copier.fit)#There're many measures calculated by lm function. 
```

### Residuals
```{r residuals}
#For example, you can look at the residuals.
#Verify the residual standard error, degree of freedom is correct.
summary(copier.fit)
res=copier.fit$residuals

SSE=sum(res^2)
df=nrow(copier)-2
MSE=SSE/df
sqrt(MSE) # Residual standard error
c(SSE,df,MSE,sqrt(MSE))
```

### Distribution of residauls 
```{r dist_res}
#Have an idea whether the residuals have a normal distribution.
plot(density(copier.fit$residuals),xlab="Residuals")

#which observaiton has the smallest residual (in absolute value)
absres=abs(res)

pos=which(absres==min(absres))
pos=which.min(absres) #an alternative way

copier.fit$fitted.values[pos]#the fitted value corresponding to this point
copier[pos,1]#the observed response of this observation
# what is residual?
```

#### Practice
```{r slr_iris}
plot(iris$Sepal.Length, iris$Petal.Length)
iris.fit <- lm(Sepal.Length~Petal.Length,data=iris)
summary(iris.fit)

## Add fitted line in the scatter plot.
## SSE, df, MSE, RSE ?
## distribution of residuals ?
```

### Beta0=0
We can force Beta0 = 0. We do this in R by telling the lm() function to not include
an intercept. We can do this with the 0 + option or the -1 option. See the examples below.
```{r omit_beta0}

lm(minutes ~ number, data=copier)
lm(minutes ~ 0 + number, data=copier)
lm(minutes ~ -1 + number, data=copier)
```

### Beta1=0
Likewise, we could use the lm() function to fit a model with no slope, but specifying the predictor variables to be constants
```{r omit_bete1}
# when Beta1 = 0.
lm(Sepal.Length ~ 1, data=iris)
```
